---
title: "R'ing with the Quran"
output: html_document
---
#Introduction
Let us use R with work with the original Arabic Quranic text. The objective of this tutorial is to get you on the road towards analyzing the Quranic text. 

The first obstacle to anyone who wants to analyze Quran -as well as any Arabic text - is the ability to deal with the Arabic text. Here I will try to make this part easier for you.

But, first let us get the Quranic text.

#Getting raw Quran

Go to http://tanzil.net/download/ and choose the right file. For this tutorial, I used "Simple Clean" without any pause marks or other options, and chose "Text(with aya numbers)", and saved it in a folder called "data". The file includes sura number followed by verse number followed by the actual text of the verse. The fields are separated by a bar "|" and has some copyright notes at the end. I had to manually remove the copyright notes before reading the file in R.


```{r}
q = read.csv("data/quran-simple-clean.txt", header=F, stringsAsFactor=F, encoding="UTF-8", sep="|")
head(q)
tail(q)
```

I hope you have the same output as mine. You might struggle to get R output correct Arabic characters as I did. I used Windows 7 and had to work with changing locale information from control panel.

So, albeit some hiccups: congrats! you have the entire Quran in R!


#Tidy up few things
Let us see the structure of our Quran in R
```{r}
str(q)
```

Since our initial file did not have headings, R automatically named the three columns as V1, V2 and V3. We need to change column names:
```{r}
colnames(q) = c("sura", "aya", "text")

```

Thats better!

With one shot lets us display the first verse in the first 10 suras
```{r}
q$text[q$aya==1 & q$sura<=10]

```

You would notice that Tanzil guys adopted the view that "بسم الله الرحمن الرحيم" is part of the first verse in all suras (expect sura no. 9)..

For me, I wanted to exclude it from all suras except suras nos 1 and 9. How to do that? 
```{r}
q$text = gsub("^بسم الله الرحمن الرحيم ","",q$text)
```

Now we are ready to go!

#Some Handy queries

##Dispaly a sura
So, let us display just a sura, say no. 100
```{r}
q[q$sura==100,]

```

##Search for a word
Here is one way to see how may times and where are the words Muhammad (محمد) is mentioned in the Quran.
```{r}
q[which(grepl("محمد", q$text)),]

```

