---
title: "First steps in Text Mining the Quran"
author: "Abdulbaqi Sharaf"
date: "Wednesday, April 29, 2015"
output: html_document
---

In the previous tutorial we created the data frame that hold the Arabic Quran called `q`
```{r}
str(q)

```
Now let us do some basic text mining works, motivated by the following document.
http://onepager.togaware.com/TextMiningO.pdf

#Text Mining Package

I wanted to experiment a bit with the `tm` package. Please install and load the package.
```{r}
library(tm)

```

The first step is to create a corpus consisting of the raw Arabic verses as `VectorSource`
```{r}
qCorpus = Corpus(VectorSource(q$text))

```

Lets `inspect` the content of this Corpus
```{r}
inspect(qCorpus[1:5])

```

We will do some more annotation work using `meta` later. For now, let us create term document matrix
```{r}
qTerms = DocumentTermMatrix(qCorpus)
qTerms
```

This produces a long matrix of documents (i.e., verses) against Quranic terms. Let us for example see a portion of this matrix by looking into documents 1 to 7 (i.e., sura Fateha) and terms say 1000 to 1005
```{r}
inspect(qTerms[1:7,1000:1005])

```

This tells us that none of these five terms appears in any of the first 7 documents. Sparsity is a known issue in document term matrices. 

#Some operations of Document Term Matrics

Lets us find some common terms in the Quran. What are terms used 100 or more times in the Quran?
```{r}
findFreqTerms(qTerms,100)

```

Interesting to see prophet Musa (Moses) "موسى" among the list.  
Note that since we did not do any stemming root words are repeated with various affixes as different words. 

Even we can create a list of most freq terms and store it in a data frame
```{r}
freq = sort(colSums(as.matrix(qTerms)),decreasing = T)
head(freq, 10)
wf = data.frame(word=names(freq), freq=freq)
```


Why not plot them using `ggplot2` package?
```{r}
library(ggplot2)
#take the most freq in a separate data frame
wfplot = subset(wf,freq>300)
ggplot(wfplot, aes(word, freq)) +
  geom_bar(stat="identity")+
  theme(axis.text.x=element_text(angle=45, hjust = 1))
```

No Wonder, Allah الله is the most frequent word. May HE be exalted!

#Word Cloud
Now let us do some more cool visualization with Word Cloud using the package `wordcloud`.Please review the package and adjust various parameters to choose the right scale and color brewer and percentage of words to rotate.

```{r}
library(wordcloud)
#I will set a seed so you can reproduce this result
set.seed(114)
wordcloud(names(freq), freq, min.freq=50, scale=c(5,.5),colors=brewer.pal(6,"Dark2"), rot.per=0.2)

```

Alayws the beatiful name ALLAH (الله) pops up in your face! Exalted be He.

#Word Length

We would like to know more about the word length in the Quran.

First lets get all words in a data frame `words` and word length in `wLen`
```{r}
words = as.matrix(colnames(qTerms))
wLen = data.frame(nletters=nchar(words))
```

Let us produce some visualization out of this.
```{r}
ggplot(wLen, aes(x=nletters))+
  geom_histogram(binwidth=1) +
  geom_vline(xintercept=mean(nchar(words)),
             colour="green", size=1, alpha=.5)+
  labs(x="Number of Letters", y="Number of Words")

```

This shows that on average word sizes are close to 5 letters. Remember we are not talking here about root words, rather raw words with all prefixes and suffixes.

#Letter frequencies 

Since we gone that far, let us conclude with analyzing the frequency of letters. First a number of packages need to be installed. 

```{r}
library(dplyr)
library(stringr)
library(qdap)
letter = str_split(words,"")
letter=sapply(letter, function(x) x[-1])
letter = unlist(letter)
letter = dist_tab(letter)
```

So, `letter` is a nice data frame that gives a list of letters with their frequency and cumulitive freq percentages. Let us produce a graph out of it
```{r}
letterMutate = mutate(letter,Letter=factor(letter$interval, levels=letter$interval[order(letter$freq)]))
ggplot(letterMutate, aes(letterMutate$Letter, weight=percent)) +
  geom_bar()+
  coord_flip()+
  ylab("Proportion")+
  xlab("Letter")+
  scale_y_continuous(breaks=seq(0,12,2),
                     label=function(x) paste0(x,"%"),
                     expand=c(0,0), limits=c(0,12))

```

#Position of letters in Quranic words

```{r}
words %>%
  lapply(function(x) sapply(letters, gregexpr, x, fixed=T)) %>%
  unlist %>%
  (function(x) x[x!=-1]) %>%
  (function(x) setNames(x, gsub("\\d", "", names(x)))) %>%
  (function(x) apply(table(data.frame(letter=names(x),
                           position=unname(x))),
                      1, function(y) y/length(x)))%>%
  qheat(high="green", low="yellow", by.column=NULL,
        values=T, digits=3, plot=F)


```

